---
title: "Fuzzy_Match"
author: "Kenneth Stevenson and Harrison Karp"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, message = FALSE)
```

Often when we want to join two tables together, we don't have unique ID to join by and have to use a column they have in common, often names. A lot of the time the data is entered manually so there might be typos in the names, or it might use a different format. For example, one might include prefixes or suffixes while the other might not. To do a normal join these columns would not match. The `fuz.merge` package allows us to join tables based on approximate matches and provides other functions to facilitate this process to ensure matches are accurate.

## The Function

The primary function of this package is `fuzzy_match`

```{r eval = FALSE, echo = TRUE}
  fuzzy_match(table1, table2, by1, by2, method = "jw", cutoff = .1, join_type = "inner", sort = "desc")
```

The function implements fuzzy matching by calculating finding the closest match between tables where the distance between the two values is less than the specified cutoff. 

## Data Sets

* The `congress` data frame has information on  541 U.S. Congressmembers and was retrieved from [data.world](https://data.world/govtrack/us-congress-legislators)

* The `politwoops` data frame has 1852 observations of various public officials who are being tracked by Politwoops for deleting tweets and was retrieved from [propublica.org](https://www.propublica.org/datastore/dataset/politicians-tracked-by-politwoops)

## Example: Inner Join

```{r}
library(fuz.merge)
congress <- dplyr::select(congress, name, gender)
politwoops <- dplyr::select(politwoops, full_name, user_name, account_type)
congress$name <- scramble(congress$name, 2)
politwoops$full_name <- scramble(politwoops$full_name, 1)

congress[1:10,"name"]
politwoops[1:10,"full_name"]

```

The scramble function randomly changed characters for the names so we can test
the fuzzy_match. First we want to use the clean() function on each dataset so we
can minimize the differences between the two columns we are going to join by.

```{r}
congress <- clean(congress, name, selected = ",", suffixes = T, prefixes = T)
congress[1:20,1]
```

So now we have removed the prefixes, suffixes and commas from each column of names.

```{r}
df <- fuzzy_match(congress, politwoops, name, full_name, cutoff = .1, 
                  join_type = "inner", sort = "desc")
df[1:15,]
```

Using the sort argument sorted the table in descending order of the match_vals 
column so the row at top is the least accurate. Looking at the by.x and by.y 
column, it's clear that this is the correct merge. The probability of having a
mismatch further down is low since the first row had the biggest string distance
between the matching columns.




